<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Setting Nodes}
-->
---
output:
  pdf_document:
    fig_caption: yes
    number_section: yes
    toc: no
  html_document:
    fig_caption: yes
    number_section: yes
    toc: yes
---

*Some notes to consider*

1. I'm not sure the division of "Using `setNode`" and "Using JAGS Code" is very clear.  We use `setNode` for everything.  We enter the JAGS code into the parameters to define the node, or we tell `setNode` to pull the values from the data.

```{r, eval=2}
install.packages("Hyde")
library(Hyde)
```

```{r}
# summary(DATASET)
```

```{r}
net <- HydeNetwork(~ x)

plot(net)
# (see `vignette("Customizing_Network_Graphs", package = "Hyde")`) 

net
```

# Root Nodes

## Binary
```{r}
net <- HydeNetwork(~ x)
net <- setNode(net, x, nodeType="dbern", p=.4)
net
```
### Using JAGS Code
I don't think there's really a JAGS code thing to do here.  The step above takes care of all of it.  

I think the only other way to specify this would be from ~xtabs (which would actually use `dcat`)

## Categorical
```{r}
net <- HydeNetwork(~ x)
net <- setNode(net, x, nodeType="dcat", pi=vectorProbs(c(3, 6, 2), x))
net
```

```{r}
net <- HydeNetwork(~ x)
net <- setNode(net, x, nodeType="dcat", 
               pi="pi.x[1] <- .2727; pi.x[2] <- .5454; pi.x[3] <- .1818")
net
```


## Normal

To define a node that has a normal distribution with $mu = 10$ and $\sigma = 5$

```{r}
net <- HydeNetwork(~ x)
net <- setNode(net, x, nodeType="dnorm", 
               mu=10, tau=1/5)
net
```

## Uniform

To define a node that has a Uniform distribution from 0 to 100:

```{r}
net <- HydeNetwork(~ x)
net <- setNode(net, x, nodeType="dunif", 
               a=0, b=100)
net
```

## Weibull


```{r}
net <- HydeNetwork(~ x)
net <- setNode(net, x, nodeType="dweib", 
               nu=1, lambda=5)
net
```

## Other types:
dgamma, dexp, dpois, dnegbin, etc



# Nodes With or Without Parents -- Manually specifying regression equations

## Nodes in the exponential family -- need regression equation + link function
### Ordinary Least Squares
```{r}
net <- HydeNetwork(~ x1 + 
                     x2 + 
                     x3 + y|x1*x2*x3)
net
plot(net)

net <- setNode(net, y, nodeType="dnorm",
               mu=fromFormula(), tau=4.8934,
               nodeFormula = y ~ 15.23 + 4.3 * x1 + 17.8 * x2 + 9.3 * x3)
net
```

Or, alternatively, one my specify the model in JAGS code in the `mu` parameter.

```{r}
net <- HydeNetwork(~ x1 + 
                     x2 + 
                     x3 + y|x1*x2*x3)
net <- setNode(net, y, nodeType="dnorm",
               mu="15.23 + 4.3 * x1 + 17.8 * x2 + 9.3 * x3", tau=4.8934)
net
```

The decision of whether to give an `R`-style formula or JAGS code is a matter of preference.  But when using `R` code, be careful that any functions used in the formula can be translated to JAGS.  A list of functions that can be translated between `R` and JAGS can be viewed by calling 

```{r, eval=FALSE}
data(jagsFunctions, package='Hyde')
jagsFunctions
```

### Lognormal
### Logistic regression nodes 
If the intercept and regression coefficients of a logistic regression are known, one may use the `ilogit` function in JAGS (inverse logit) to back transform the predicted response to a normally distributed probability.
```{r}
net <- HydeNetwork(~ x1 + 
                     x2 + 
                     x3 + y|x1*x2*x3)

net <- setNode(net, y, nodeType="dnorm",
               mu="ilogit(15.23 + 4.3 * x1 + 17.8 * x2 + 9.3 * x3)", tau=4.8934)
net

writeNetworkModel(net, TRUE)
```

If desired, one could also format this as a bernoulli distributed random variable.
```{r}
net <- HydeNetwork(~ x1 + 
                     x2 + 
                     x3 + y|x1*x2*x3)

net <- setNode(net, y, nodeType="dbern",
               p="ilogit(15.23 + 4.3 * x1 + 17.8 * x2 + 9.3 * x3)")
net
```

### Poisson Regression nodes
### Gamma regression nodes

## Categorical Nodes -- list (?) of multinomial logistic regression equations -- this will be tricky

## Continuous Nodes
### Regression equation with Gaussian errors

## Manual Specification using JAGS code
### Linear regression equation with cauchy errors
### Mixture distributions (do this last!)
#### Zero-inflated Poisson
#### Zero inflated negative binomial




# Node With or Without Parents -- Reading formula from existing model objects

## Reading from `lm` objects
## Reading from `glm` objects - can we make this work for lognormal
## Reading from multinom objects
## Reading from survreg objects (note the list component modelObject$dist that stores the type of distribution used for the parametric survival regression model)



# Learning Node Distributions from Training Data

## Default behavior of `HydeNetwork(.formula, data=trainingData)`
## Tutorial involving user-specified distributions for particular nodes and estimation of regression equations from data 
one node may be survival (with censoring), one may be binary, one may be normal, one may be Poisson, and the other may be categorical




```{r, echo=FALSE}
#----------------------------------------------------------------------------------------------------------
# 1) Root Nodes
#----------------------------------------------------------------------------------------------------------
#  1.1) Binary
#    1.1.1) setNode(HydeNetwork, x, nodeType="dbern", p=0.62)   (need logic?  0<=p<=1)
#    1.1.2) Using JAGS code (nodeType="dbern", params="pi=0.28;"..or whatever the correct syntax is)
#  1.2) Categorical
#    1.2.1) setNode(HydeNetwork, x, nodeType="dcat",  p=c(0.62,0.08))
#           (need logic? if(length(p)==(K-1)) 0<=sum(p)<=1 else if(length(p)==K) sum(p)==1...)
#    1.2.2) Using JAGS code into the params argument, i.e., nodeType="dcat", 
#           params="pi.wells[1] = .85; pi.wells[2] = .12; pi.wells[3] = .03"
#           need specifics as to how that string needs to be constructed - for instance, 
#           do you need to use the name "pi.wells"?
#  1.3) Normal
#    1.3.1) setNode(HydeNetwork, x, nodeType="dnorm", mu=5, sigma=1.4)
#    1.3.2) Using JAGS code
#  1.4 Uniform
#    1.4.1) setNode(HydeNetwork, x, nodeType="dunif", a=0, b=10)
#    1.4.2) Using JAGS code
#  1.5 Weibull
#    1.5.1) setNode(HydeNetwork, x, nodeType="dweib", nu=3, lambda=1.5)
#    1.5.2) Using JAGS code
#  1.6 -- 1.1x  [dgamma, dexp, dpois, dnegbin, etc...]

#----------------------------------------------------------------------------------------------------------
# 2) Nodes With or Without Parents -- Manually specifying regression equations
#    >>> Do we need error checking to ensure alignment between stated terms in regression equations and 
#        network structure (i.e., if X1 and X2 are in the regression equation, they need to be stored
#        as parents of Y in the HydeNetwork object)?
#----------------------------------------------------------------------------------------------------------
#  2.1) Nodes in the exponential family -- need regression equation + link function
#    2.1.1) Ordinary Least Squares
#    2.1.2) Lognormal 
#    2.1.3) Logistic Regression nodes ... I think JAGS has a logit() function that may be useful
#    2.1.4) Poisson Regression nodes 
#    2.1.5) Gamma regression
#  2.2) Categorical nodes -- list (?) of multinomial logistic regression equations -- this will be tricky.
#  2.3) Continuous nodes
#    2.3.1) Regression equation with Gaussian errors
#  2.4) Manual specification using JAGS code
#    2.4.1) Linear regression equation with cauchy errors
#    2.4.2) Mixture distributions (do this last!)
#      2.4.2.1) Zero-inflated poisson
#      2.4.2.2) Zero-inflated negative binomial

#----------------------------------------------------------------------------------------------------------
# 3) Nodes With or Without Parents -- Reading formula from existing model objects
#    >>> Do we need error checking to ensure alignment between stated terms in regression equations and 
#        network structure (i.e., if X1 and X2 are in the regression equation, they need to be stored
#        as parents of Y in the HydeNetwork object)?
#----------------------------------------------------------------------------------------------------------
#  3.1) Reading from lm objects
#  3.2) Reading from glm objects  - can we make this work for lognormal?
#  3.3) Reading from multinom objects
#  3.4) Reading from survreg objects (note the list component modelObject$dist that stores the type
#       of distribution used for the parametric survival regression model)


#----------------------------------------------------------------------------------------------------------
# 4) Learning Node Distributions from Training Data
#----------------------------------------------------------------------------------------------------------
#  4.1) Default behavior of HydeNetwork(.formula, data=trainingData)
#  4.2) Tutorial involving user-specified distributions for particular nodes and estimation of regression
#       equations from data -- one node may be survival (with censoring), one may be binary, one may be normal,
#       one may be Poisson, and the other may be categorical
```
